Estimating Correspondences of Deformable Objects ``In-the-wild"

### 2015-11-06 contents/evaluation.tex
Save

\subsubsection{Databases \& Error Metrics}
For face experiment, both models were built using the 811 training images of the Labelled
Faces Parts in-the-Wild (LFPW) \cite{Asthana2014} database. Classic AAM were built from 68 point landmark annotations using the standard procedure described in \cite{Cootes2001,Matthews2004}; piecewise affine was used as their motion model. Our dAAMs were built as the pipeline proposed. Results for this experiment are reported over the 224 testing images of the LFPW database using again 68 point ground truth landmark annotations. For this experiment, we report results using pixels intensities (INT) for both models.

@@

\subsubsection{Experiments}

@@

CED Curves for this experiment are shown in Figure \ref{fig:daam_benchmark}. We can readily observe that the accuracy achieved by ear models, in terms of the previous normalized point-to-point error measure, is lower than the one achieved by face models. This result is expected due to the more complex structure of the ear shape. Through visual inspection we determined that fitting errors below 0.1 for ears typically look plausible under the previous error measure where fitting errors below 0.6 for faces. Results show that, for this experiment, the fitting accuracy achieved by our dAAMs is marginally better to the one achieved by classic AAMs using intensity representations. In fact, classic AAM fail to improve the accuracy of the original initialisation on ears. Therefore the propose pipeline is capable of dealing with the complex structure of non-rigid shapes and learn dAAMs from simple curve line annotations that can compete and even surpass (in the case of a simple pixel intensity image representation) classic AAM, build from carefully annotated point landmarks, is quite remarkable.

@@

\subsubsection{Dataset \& Error Metric}
There are significant numbers of datasets exist for human pose estimation, where FLIC\cite{sapp2013modec}, BBC Pose\cite{pfister2015flowing}, Fashion Pose\cite{dantone2013human} and MPII\cite{andriluka14cvpr} attracted most public attention. In our work, we reported our performance comparison on BBC Pose Dataset\cite{pfister2015flowing}, which has relatively most consistent skeleton annotation among the others.

@@

Two experiments are undertaken in this section: 1) presenting training classic patch-based AAM from outline annotation provides remarkable joints estimation; 2) supporting outline annotation by comparing with training classic patch-based AAM from joint annotation.

First experiment presented the accuracy of predicting joints positions from establishing dense correspondences with outline fitting. We build patch-based AAM using 891 training hand images from a combination of datasets, where each image has 29 landmarks annotated alone its boundary. SIFT \cite{PoseletsICCV09} feature is used from image representation for our model. Model fitting is initialised based on current state-of-the-art human pose estimation algorithms on BBC Pose dataset. Results for this experiment are reported over 1000 testing images BBC Pose Provided, which human upper-body pose are landmark with 7 points on skeleton. For our model, joints are projected from the dense correspondence generated from our proposed method.

@@

\begin{figure*}[!t]
    \newcommand{\fh}{0.24\columnwidth}
    \centering
    \includegraphics[height=\fh]{resources/Fixing/fix_3}
    \includegraphics[height=\fh]{resources/Fixing/fix_5}
    \includegraphics[height=\fh]{resources/Fixing/fix_6}
    \includegraphics[height=\fh]{resources/Fixing/fix_7}
    \includegraphics[height=\fh]{resources/Fixing/fix_8}
    \includegraphics[height=\fh]{resources/Fixing/fix_9}
    \includegraphics[height=\fh]{resources/Fixing/fix_10}
    \includegraphics[height=\fh]{resources/Fixing/fix_12}
    \includegraphics[height=\fh]{resources/Fixing/fix_13}
    \includegraphics[height=\fh]{resources/Fixing/fix_14}
    \includegraphics[height=\fh]{resources/Fixing/fix_15}
    \includegraphics[height=\fh]{resources/Fixing/fix_17}
    \includegraphics[height=\fh]{resources/Fixing/fix_18}
    \includegraphics[height=\fh]{resources/Fixing/fix_19}
    \caption{Demonstration of annotation correction using our method for experiment \ref{exp:qualitative}. \emph{Red} dots refer to officially provided landmarks, and \emph{green} dots are corrected position.}
    \label{fig:qualitative}
\end{figure*}

\begin{figure*}[!t]
    \newcommand{\ofh}{0.24\columnwidth}
    \centering
    \includegraphics[height=\ofh]{resources/Fittings/1.eps}
    \includegraphics[height=\ofh]{resources/Fittings/3.eps}
    \includegraphics[height=\ofh]{resources/Fittings/19.eps}
    \includegraphics[height=\ofh]{resources/Fittings/4.eps}
    \includegraphics[height=\ofh]{resources/Fittings/6.eps}
    \includegraphics[height=\ofh]{resources/Fittings/15.eps}
    \includegraphics[height=\ofh]{resources/Fittings/7.eps}
    \includegraphics[height=\ofh]{resources/Fittings/8.eps}
    \includegraphics[height=\ofh]{resources/Fittings/13.eps}
    \includegraphics[height=\ofh]{resources/Fittings/17.eps}
    \\
    \includegraphics[height=\ofh]{resources/Fittings/21.eps}
    \includegraphics[height=\ofh]{resources/Fittings/23.eps}
    \includegraphics[height=\ofh]{resources/Fittings/25.eps}
    \includegraphics[height=\ofh]{resources/Fittings/27.eps}
    \includegraphics[height=\ofh]{resources/Fittings/29.eps}
    \includegraphics[height=\ofh]{resources/Fittings/31.eps}
    \includegraphics[height=\ofh]{resources/Fittings/33.eps}
    \includegraphics[height=\ofh]{resources/Fittings/35.eps}
    \includegraphics[height=\ofh]{resources/Fittings/36.eps}
    \includegraphics[height=\ofh]{resources/Fittings/37.eps}
    \includegraphics[height=\ofh]{resources/Fittings/38.eps}
    \includegraphics[height=\ofh]{resources/Fittings/39.eps}
    \caption{Demonstration of outline fitting of patch-based AAM on arms.}
    \label{fig:outline_fitting}
\end{figure*}

@@

Figure \ref{exp:qualitative} presents corrections of joint annotations using patch-based SDM trained as experiment \ref{exp:benchmark}. \emph{Red} dots shows ``ground truth" points while \emph{green} dots shows landmarks after correction using our methodology. There is no doubt that points after correction show more consistency among images. Images are croped to hand only for better visualisation.

>>>>

\textbf{Databases \& Error Metrics}
For face experiment, both models were built using the 811 training images of the Labelled
Faces Parts in-the-Wild (LFPW) \cite{Asthana2014} database. Classic AAM were built from 68 point landmark annotations using the standard procedure described in \cite{Cootes2001,Matthews2004}; piecewise affine was used as their motion model. Our dAAMs were built as the pipeline proposed. Results for this experiment are reported over the 224 testing images of the LFPW database using again 68 point ground truth landmark annotations. For this experiment, we report results using pixels intensities (INT) for both models.

@@

@@

\textbf{Experiments}
CED Curves for this experiment are shown in Figure \ref{fig:daam_benchmark}. We can readily observe that the accuracy achieved by ear models, in terms of the previous normalised point-to-point error measure, is lower than the one achieved by face models. This result is expected due to the more complex structure of the ear shape. Through visual inspection we determined that fitting errors below 0.1 for ears typically look plausible under the previous error measure where fitting errors below 0.6 for faces. Results show that, for this experiment, the fitting accuracy achieved by our dAAMs is marginally better to the one achieved by classic AAMs using intensity representations. In fact, classic AAM fail to improve the accuracy of the original initialisation on ears. Therefore the propose pipeline is capable of dealing with the complex structure of non-rigid shapes and learn dAAMs from simple curve line annotations that can compete and even surpass (in the case of a simple pixel intensity image representation) classic AAM, build from carefully annotated point landmarks, is quite remarkable.

@@

\textbf{Dataset \& Error Metric}
There are significant numbers of datasets exist for human pose estimation, where FLIC\cite{sapp2013modec}, BBC Pose\cite{pfister2015flowing}, Fashion Pose\cite{dantone2013human} and MPII\cite{andriluka14cvpr} attracted most public attention. In our work, we reported our performance comparison on BBC Pose Dataset\cite{pfister2015flowing}, which has relatively most consistent skeleton annotation among the others.

@@

\textbf{Experiments} 
Two experiments are undertaken in this section: 1) presenting training classic patch-based AAM from outline annotation provides remarkable joints estimation; 2) supporting outline annotation by comparing with training classic patch-based AAM from joint annotation.

First experiment presented the accuracy of predicting joints positions from establishing dense correspondences with outline fitting. We build patch-based AAM using 891 training hand images from a combination of datasets, where each image has 29 landmarks annotated alone its boundary. SIFT \cite{PoseletsICCV09} feature is used from image representation for our model. Model fitting is initialised based on current state-of-the-art human pose estimation algorithms (including Buehler \cite{buehler2011upper}, Charles14 \cite{charles2014upper}, Charles13 \cite{charles2013domain}, Pfister14 \cite{pfister2015deep}, Ramanan \cite{yang2013articulated} and Pfister15 \cite{pfister2015flowing}) on BBC Pose \cite{pfister2015flowing} dataset. Results for this experiment are reported over 1000 testing images BBC Pose Provided, which human upper-body pose are landmark with 7 points on skeleton. For our model, joints are projected from the dense correspondence generated from our proposed method.

@@

\begin{figure*}[!t]
    \newcommand{\fh}{0.24\columnwidth}
    \centering
    \includegraphics[height=\fh]{resources/Fixing/fix_1}
    \includegraphics[height=\fh]{resources/Fixing/fix_2}
    \includegraphics[height=\fh]{resources/Fixing/fix_3}
    \includegraphics[height=\fh]{resources/Fixing/fix_5}
    \includegraphics[height=\fh]{resources/Fixing/fix_6}
    \includegraphics[height=\fh]{resources/Fixing/fix_7}
    \includegraphics[height=\fh]{resources/Fixing/fix_8}
    \includegraphics[height=\fh]{resources/Fixing/fix_9}
    \includegraphics[height=\fh]{resources/Fixing/fix_10}
    \includegraphics[height=\fh]{resources/Fixing/fix_20}
    \includegraphics[height=\fh]{resources/Fixing/fix_12}
    \includegraphics[height=\fh]{resources/Fixing/fix_13}
    \includegraphics[height=\fh]{resources/Fixing/fix_14}
    \includegraphics[height=\fh]{resources/Fixing/fix_15}
    \includegraphics[height=\fh]{resources/Fixing/fix_17}
    \includegraphics[height=\fh]{resources/Fixing/fix_18}
    % \includegraphics[height=\fh]{resources/Fixing/fix_19}
    \caption{Demonstration of annotation correction using our method for experiment \ref{exp:qualitative}. \emph{Red} dots refer to officially provided landmarks, and \emph{green} dots are corrected position.}
    \label{fig:qualitative}
\end{figure*}

\begin{figure*}[!t]
    \newcommand{\ofh}{0.24\columnwidth}
    \centering
    \includegraphics[height=\ofh]{resources/Fittings/20.eps}
    \includegraphics[height=\ofh]{resources/Fittings/3.eps}
    \includegraphics[height=\ofh]{resources/Fittings/19.eps}
    \includegraphics[height=\ofh]{resources/Fittings/4.eps}
    \includegraphics[height=\ofh]{resources/Fittings/6.eps}
    \includegraphics[height=\ofh]{resources/Fittings/15.eps}
    \includegraphics[height=\ofh]{resources/Fittings/7.eps}
    \includegraphics[height=\ofh]{resources/Fittings/8.eps}
    \includegraphics[height=\ofh]{resources/Fittings/13.eps}
    \includegraphics[height=\ofh]{resources/Fittings/17.eps}
    \\
    \includegraphics[height=\ofh]{resources/Fittings/21.eps}
    \includegraphics[height=\ofh]{resources/Fittings/23.eps}
    \includegraphics[height=\ofh]{resources/Fittings/25.eps}
    \includegraphics[height=\ofh]{resources/Fittings/27.eps}
    \includegraphics[height=\ofh]{resources/Fittings/29.eps}
    \includegraphics[height=\ofh]{resources/Fittings/31.eps}
    \includegraphics[height=\ofh]{resources/Fittings/33.eps}
    \includegraphics[height=\ofh]{resources/Fittings/35.eps}
    \includegraphics[height=\ofh]{resources/Fittings/36.eps}
    \includegraphics[height=\ofh]{resources/Fittings/37.eps}
    \includegraphics[height=\ofh]{resources/Fittings/38.eps}
    \includegraphics[height=\ofh]{resources/Fittings/39.eps}
    \caption{Demonstration of outline fitting of patch-based AAM on arms.}
    \label{fig:outline_fitting}
\end{figure*}

@@

Figure \ref{exp:qualitative} presents corrections of joint annotations using patch-based SDM trained as experiment \ref{exp:benchmark}. \emph{Red} dots shows ``ground truth" points while \emph{green} dots shows landmarks after correction using our methodology. There is no doubt that points after correction show more consistency among images. Images are cropped to hand only for better visualisation.

### 2015-11-06 contents/implementation.tex
Save

In order to fully capture the variability among most deformable objects' shapes annotations, we use a representation based on Support Vector Shapes (SVS) \cite{Nguyen2013}. An SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a)~the representation is completely generic, e.g. it can be applied to sparse landmark points, curves lines or a combination of the two, and~(b) it fuses inconsistent landmarks into consistent and directly comparable decision functions. Furthermore, this representations is also robust against noise, missing data and outliers \cite{Nguyen2013}.

@@

After constructing the SVS representation for all images, the next step is to apply a simple similarity alignment over them. This is done because the goal here is to build a model capable of effectively representing non-rigid local shape deformations rather than global rotation, translation and scaling. The alignment is performed by using the Iterative Closest Point (ICP) algorithm \cite{Besl1992} on the annotated landmarks point cloud of the training images.

@@

Instead of the \emph{motion} basis used in multiframe optical flow formulation of \cite{Garg:2013hu}, we build a \emph{correspondence} basis that introduces constraints on how points of different shapes are matched to each other. Every pixel of the reference SVS image is matched to its corresponding position at every training SVS image and in this way defines a \emph{correspondence vector}. This vector consists of the 2D locations of the specific point in all SVS images. To form this vector, the training images are arranged in an arbitrary order. Similarly to the order of training samples when Principal Component Analysis (PCA) is applied, this order does not affect the result of our method and any re-ordering would produce exactly the same results.

@@

To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply Non-rigid Iterative Closest Point (NICP)\cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. Finally, the correspondence basis is found by applying PCA on these  correspondence vectors and keeping only the first $R$ principal components.

>>>>

In order to fully capture the variability among most deformable objects' shapes annotations, we use a representation based on SVS \cite{Nguyen2013}. An SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a)~the representation is completely generic, e.g. it can be applied to sparse landmark points, curves lines or a combination of the two, and~(b) it fuses inconsistent landmarks into consistent and directly comparable decision functions. Furthermore, this representations is also robust against noise, missing data and outliers \cite{Nguyen2013}.

@@

After constructing the SVS representation for all images, the next step is to apply a simple similarity alignment over them. This is done because the goal here is to build a model capable of effectively representing non-rigid local shape deformations rather than global rotation, translation and scaling. The alignment is performed by using the ICP algorithm \cite{Besl1992} on the annotated landmarks point cloud of the training images.

@@

Instead of the \emph{motion} basis used in multiframe optical flow formulation of \cite{Garg:2013hu}, we build a \emph{correspondence} basis that introduces constraints on how points of different shapes are matched to each other. Every pixel of the reference SVS image is matched to its corresponding position at every training SVS image and in this way defines a \emph{correspondence vector}. This vector consists of the 2D locations of the specific point in all SVS images. To form this vector, the training images are arranged in an arbitrary order. Similarly to the order of training samples when PCA is applied, this order does not affect the result of our method and any re-ordering would produce exactly the same results.

@@

To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply NICP \cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. Finally, the correspondence basis is found by applying PCA on these  correspondence vectors and keeping only the first $R$ principal components.
